Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-10-28 manuscript.tex
latest

\subsection*{Grime's triangle}\label{sec:grime}

@@

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$ (which is already small). In plants and corals, the ability to secure territories $c$ is strongly associated with rapid individual growth, and rapid individual growth is a defining feature of the competitor trait cluster \cite{grime_1977,darling_2012}. Evolution for higher $b$ (which is closely related to the intrinsic low-density growth rate $r=b-d$ ) is not usually thought to be the expected outcome in high-density competitive environments. We will discuss this further below in the context of MacArthur's $r$/$K$ scheme. For now we simply note that the analogue of Grime's ideal environment for fishes occurs when resources limitations are not severe and the environment varies predictably over large temporal and spatial scales \cite{winemiller_1992,winemiller_2015}. The associated ``periodic'' strategy is characterized by enormous brood sizes, suggesting a strategy of ensuring that more propagules actually end up contesting areas favorable for development (higher $b$).

@@

To calculate the relative fluctuations in $\sum_{j\neq i} c_j x_j$, we use a similar approximation as for $\Delta_r n_i$: $\tilde{p}$ is approximated by the ${\mathbf x}$ dispersal probabilities in a territory where at least two focal propagule is assumed to be present. All covariances are now zero, so that $\sigma^2(\sum c_j x_j)=\sum c_j^2 \sigma^2(x_j)$, where $\sigma^2(x_j)=l_j$ for $j\neq i$. The expression for $\sigma^2(x_i)$ is more complicated. We assume $p(x_i=0)\approx 0$ without loss of generality (since otherwise $D\gg 1$ and $\Delta n_a$ is negligible). Then
\begin{equation}
\sigma^2(x_i)=\frac{l_i^2}{D}\left(1-\frac{1}{D}\right)+\frac{l_i}{D},
\end{equation}
where $D=1-(1+l_i)e^{-l_i}$, analogous to Eq. \eqref{eq:varr}, and 
\begin{equation}
\frac{\sigma(\sum c_j x_j)}{\langle\sum c_j x_j\rangle} \approx\frac{\left(\sum_{j\neq i} c_j^2 l_j + c_i^2 \sigma^2(x_i)\right)^{1/2}}{\sum_{j\neq i} c_j l_j + c_i l_i/D} \label{eq:cva}.
\end{equation}

>>>>

\subsection*{Primary strategies and Grime's triangle}

@@

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$. In plants and corals, the ability to secure territories $c$ is strongly associated with rapid individual growth; not surprisingly, rapid individual growth is a defining feature of the competitor trait cluster \cite{grime_1977,darling_2012}. The analogue of Grime's ideal environment for fishes occurs when resources limitations are not severe and the environment varies predictably over large temporal and spatial scales \cite{winemiller_1992,winemiller_2015}. The associated ``periodic'' strategy is characterized by enormous brood sizes, suggesting a strategy of ensuring that more propagules actually end up contesting areas favorable for development (higher $b$). Evolution for higher $b$ (which is closely related to the intrinsic low-density growth rate $r=b-d$) will be discussed further in the Discussion in the context of MacArthur's $r$/$K$ scheme.

@@

To calculate the relative fluctuations in $\sum_{j\neq i} c_j x_j$, we use a similar approximation as for $\Delta_r n_i$: $\hat{p}$ is approximated by the ${\mathbf x}$ dispersal probabilities in a territory where at least two focal propagule is assumed to be present. All covariances are now zero, so that $\sigma^2(\sum c_j x_j)=\sum c_j^2 \sigma^2(x_j)$, where $\sigma^2(x_j)=l_j$ for $j\neq i$. The expression for $\sigma^2(x_i)$ is more complicated. We assume $p(x_i=0)\approx 0$ without loss of generality (since otherwise $D\gg 1$ and $\Delta n_a$ is negligible). Then
\begin{equation}
\sigma^2(x_i)=\frac{l_i^2}{D}\left(1-\frac{1}{D}\right)+\frac{l_i}{D},
\end{equation}
where $D=1-(1+l_i)e^{-l_i}$, analogous to Eq. \eqref{eq:varr}, and 
\begin{equation}
\frac{\sigma(\sum c_j x_j)}{\langle\sum c_j x_j\rangle} \approx\frac{\left(\sum_{j\neq i} c_j^2 l_j + c_i^2 \sigma^2(x_i)\right)^{1/2}}{\sum_{j\neq i} c_j l_j + c_i l_i/D} \label{eq:cva}.
\end{equation}
